Hilbert space methods to approximate Gaussian processes using Stan

### 2019-09-07 Paper/manuscript/manuscript_06.tex
improved figures and added all study cases except birthday dataset

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

@@

Although there are several GP specific software packages available to date (GPML, GPstuff, GPy, GPflow, cite them!), each provide efficient implementations only for a restricted range of GP based models. In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.

@@

The mean and covariance functions completely characterize the Gaussian process prior, and control the a priori behavior of the function $f$. Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution
$\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$,
 where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$. The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
The joint distribution of $\bm{f}$ and a new $f^*$ is also a multivariate Gaussian as,
%
\begin{align*}
p(\bm{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
\bm{f} \\ 
f^*
\end{array} \right] \,\middle|\, \bm{0},\left[ \begin{array}{cc}
\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},f^*} \\ 
\bm{k}_{f^*,\bm{f}} & k_{f^*,f^*}
\end{array} \right] \right),
\end{align*}

@@

The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs. A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations. Isotropic covariance functions are those that are function of the distance between observations, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant. The most commonly used distance between observations is the norm L2 $(|\bm{x}-\bm{x}'|_{L2})$, also known as Euclidean distance, although other types of distances can be considered.

@@

Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$. Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have 
%
\begin{align}
\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenvalue_multi} \\
%
\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenfunction_multi}
\end{align}

@@

(Gabi: The following paragraph is a first idea to introduce that the recommendations and diagnostics depends on the kernel considered. And that, at the moment, we have built these recommendations only for the square exponential kernel. I know that what I wrote in not enough and should be a bit more elaborate; your suggestions where and how elaborate it would be useful.)

Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function. In this work we built these relationships for the square exponential covariance function. For other kernels, the relationships will be slightly different.

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig1_legend.pdf}}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
  \label{fig1_Post_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig2_legend.pdf}}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
  \label{fig2_Post_L}
\end{figure}

@@

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\ 
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
\end{tabular}
\end{center}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig3_Post_part1}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model. (left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$. (right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig4_MSE_vs_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig5_legend.pdf}}
\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig5_lscale_vs_J}
\end{figure}

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figure \ref{fig6_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to 
% the performance of the HSGP approximation.
a close approximation of the HSGP model, in the case of a GP with square exponential covariance function and a single input dimension. (Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?). More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}). We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
%
\begin{eqnarray*}
 \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
\end{eqnarray*}

@@

As stated above, Figure \ref{fig6_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c}. If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$. Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
\hspace{3mm}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function. The right-side plot is a zoom in of the left-side plot.}
  \label{fig6_lscale_vs_J_vs_c}
\end{figure}

@@

Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending on three variables, $m$, $c$ and $\ell$. An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented. However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.

(Gabi: The following paragraph is another idea to introduce that the diagnosis plot in Figure \ref{fig6_lscale_vs_J_vs_c} depends on the kernel considered. At the moment we have built this figure only for the square exponential kernel.)

Ultimately, Figure \ref{fig6_lscale_vs_J_vs_c} showing the relationships among the number of basis functions, the boundary factor and the lengthscale of the function depends on the particular choice of covariance function. In this work we have built these relationships only for the square exponential covariance function. For other kernels, the Figure will be slightly different.

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
\vspace{-2mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
\vspace{-3mm}
\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
  \label{fig7_posterior_varing_lscale_part1}
\end{figure}

>>>>

\usepackage{subfigure}
\usepackage{multirow,slashbox}
\usepackage{array}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

@@

Although there are several GP specific software packages available to date (GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP based models. In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.

@@

The mean and covariance functions completely characterize the Gaussian process prior, and control the a priori behavior of the function $f$. Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution
$\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$,
 where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$, which needs be positive semidefinite. The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
The joint distribution of $\bm{f}$ and a new $f^*$ is also a multivariate Gaussian as,
%
\begin{align*}
p(\bm{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
\bm{f} \\ 
f^*
\end{array} \right] \,\middle|\, \bm{0},\left[ \begin{array}{cc}
\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},f^*} \\ 
\bm{k}_{f^*,\bm{f}} & k_{f^*,f^*}
\end{array} \right] \right),
\end{align*}

@@

The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs. A covariance function needs be symmetric and positive semidefinite. A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations. Isotropic covariance functions are those that are function of the distance between observations, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant. The most commonly used distance between observations is the norm L2 $(|\bm{x}-\bm{x}'|_{L2})$, also known as Euclidean distance, although other types of distances can be considered.

@@

Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$. Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have 
%
\begin{align}
\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenfunction_multi} \\
%
\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenvalue_multi}
\end{align}

@@

Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function. In this work we built these relationships for the square exponential covariance function and Matern ($\nu=3/2$) covariance function. For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects.

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 5mm 0mm 5mm 0mm, clip]{fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 122mm 10mm, clip]{fig1_legend.pdf}}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
  \label{fig1_Post_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 5mm 0mm 5mm 0mm, clip]{fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 120mm 0mm, clip]{fig2_legend.pdf}}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
  \label{fig2_Post_L}
\end{figure}

@@

\begin{figure}
\begin{center}
\begin{tabular}{| c | c | c | c |}
\hline
c  = 1.05 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part1.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part1.pdf} & \multirow{36}{1.3cm}{\includegraphics[scale=0.35, trim = 28mm 40mm 100mm 30mm, clip]{fig3_legend.pdf}}\\ 
\cline{1-3}
c  = 1.1 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part2.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part2.pdf} &\\
\cline{1-3}
c  = 1.2 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part3.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part3.pdf} &\\
\cline{1-3}
c  = 1.5 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part4.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part4.pdf} & \\
\cline{1-3}
c  = 2 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part5.pdf} &  \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part5.pdf} &\\
\cline{1-3}
c  = 2.5 & \includegraphics[scale=0.23, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Post_part6.pdf} & \includegraphics[scale=0.23, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Cov_part6.pdf} &\\
\hline
\end{tabular}
\end{center}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig3_Post_part1}
\end{figure}

%\begin{figure}
%\begin{tabular}{ c c c c c}
%\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} &  \hspace{-0mm}
%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\ 
%\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} & \hspace{-0mm}
%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.35, trim = 25mm 63mm 120mm 50mm, clip]{fig3_legend.pdf}\\
%\includegraphics[scale=0.20, trim = 10mm 5mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} & \hspace{-0mm}
%\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} 
%\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} & 
%\end{tabular}
%\caption{Mean posterior predictive functions (left-first and right-first columns) and covariance functions (left-second and right-second columns) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
%  \label{fig3_Post_part1}
%\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 0mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model. (left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$. (right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig4_MSE_vs_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 28mm 28mm 119mm 0mm, clip]{fig5_legend.pdf}}
\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig5_lscale_vs_J}
\end{figure}

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to 
% the performance of the HSGP approximation.
a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension. More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}). Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern ($\nu=3/2$) covariance function. We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
%
\begin{eqnarray*}
 \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
\end{eqnarray*}

@@

As stated above, Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c}. If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased. We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$. Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
\hspace{3mm}
\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function. The right-side plot is a zoom in of the left-side plot.}
  \label{fig6_lscale_vs_J_vs_c}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_Matern.pdf}}
\hspace{3mm}
\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\frac{3}{2}$) covariance function. The right-side plot is a zoom in of the left-side plot.}
  \label{fig6_lscale_vs_J_vs_c_Matern}
\end{figure}

@@

Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, and result in a surface depending on three variables, $m$, $c$ and $\ell$. An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented. However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
\begin{center}
\end{center}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4, trim = 10mm 100mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 10mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
\vspace{-3mm}
\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
  \label{fig7_posterior_varing_lscale_part1}
\end{figure}
